Syntax-Based Concept Alignment for Domain-Specific Machine Translation

### 2021-05-07 paper/paper.tex
summarized pos-equiv

\subsubsection{Alignment criteria} \label{criteria}
% label matching
% pos equivalence
% known divergence -> cite Dorr, few examples
% known alignment -> SMT integration

>>>>

\subsubsection{Alignment criteria} \label{criteria}
While alignment criteria are customizable, to better understand the extraction algorithm described in the above we dedicate this section to describing the ones that are currently used by default.

\paragraph{Matching UD labels}
The most obvious, but also most effective idea is to determine alignability based on comparing the dependency labels of the candidate UD tree pair. 
In particular, according to this idea, two subtrees, in \textit{matching context}, i.e. attached to aligned heads constitute an alignment if their roots share the same dependency label. 
Intuitively, this means that they are in the same relation with their heads.

Note that, since the root of US trees is always attached to a a fake node with an arc labelled \texttt{root}, this criterion also makes it so that full sentences are always considered to align with each other, something which is desirable since we assume the parallel texts that are fed to our program to be sentence-aligned.

\paragraph{Part-Of-Speech equivalence}
CoNNL-U files provide information not only on the syntactic role of the tokens a sentence is composed of, but also on their grammatical categories, represented as Universal Part-Of-Speech (POS) tags. 
Intuitively, if the nodes of two trees in matching contexts have the same POS tags, the two trees are more likely to correspond to each other than if not.

This is especially true if focus, for instance, solely on the open class words (in the current implementation, defined as in the UD documentation \cite{TODO:}), thus ignoring prepositions, determiners, auxiliary verbs and all other function words, which tend to behave differently across different languages.

As a consequence, a useful relation to define between dependency trees is that of \textit{POS-equivalence}: two dependency trees $t$, $u$ are POS-equivalent if $M_1 = M_2 \neq \emptyset$, where $M_i$ is defined as the multiset of POS tags of all the open class word nodes of $t_i$.

Applied alone, this criterion can be used to capture correspondences that would be missed by only matching UD labels, thus increasing recall, but a decrease in precision is also to be expected. 
However, since alignment criteria are defined as boolean functions, it is easy to combine them so to apply them simultaneously: combining these first two criteria, for instance, is useful in cases where precision is more important than recall.

\paragraph{Known translation divergence}
% Dorr

\paragraph{Known alignment}
% example application: SMT integration
